Attrition in Online Learning: Understanding Persistence and Dropout in Massive Open Online Courses

### 2014-10-28 paper_v1.Rnw
abstract and restructuring

\begin{abstract}
% Persistence and attrition in MOOCs are systematically analyzed using self-report and behavioral data collected from 20879 online learners in 18 courses. Study 1 offers insights into reasons for disengaging from MOOCs and explores relationships with prior behavior and reported intentions. Study 2 is a case study to develop a deeper understanding of attrition in MOOCs by conducting a case study. Targeting online learners who were predicted likely dropouts in a particular course were invited to provide feedback via a survey. (abstract needs work...)
Abstract goes here.
\end{abstract}

@@

Educational environments have become increasingly diverse. Traditional schools and universities have a characteristically rigid structure, including instructor-defined---even nationally agreed---syllabi, fixed time schedules, entry requirements, and material costs to enter and exit. Novel institutional structures have been developed to overcome particular constraints. Community colleges, for instance, were created in an attempt to democratize education by offering instruction at a lower cost and by accommodating people with less flexible schedules \cite{goldrick2010challenges}. Distance learning programs intended to deliver education in remote parts of the world and for people who simply could not attend in-person classes. Course materials, including assessments, were delivered through mail (correspondence education), radio, television, and eventually the Internet, thereby addressing geographical and time related constraints of traditional instruction \cite{moore1996distance}.

@@

\begin{description}
  \item[RQ1] How do prior learner characteristics (demographics, geographic location, intentions and motivations, and previous online courses) influence learners' likelihood of disengaging from a MOOC?
  \item[H1] Successful MOOC learners exhibit higher levels of goal striving than unsuccessful ones.
\end{description}

@@

\begin{description}
  \item[RQ2] What proportion of learners who disengaged were satisfied with their achievements in the course?
  \item[H3] Successful MOOC learners have a more fluid conception of intelligence (i.e., more of a growth than fixed mindset) than unsuccessful ones.
\end{description}

@@

\begin{description}
\item[RQ3a] What reasons do learners report for disengaging from a MOOC?
\item[RQ3b] How do reasons for disengaging vary across prior learner characteristics (cf. RQ1)?
\item[RQ4] How does behavior in MOOCs vary by reasons for disengaging?
\end{description}

@@

Seventeen MOOCs on topics in a variety of disciplines were selected for investigation (see Table \ref{tab:s1sum}). For each course, we merged data from the initial course survey (distributed in the first weeks of the course), the feedback surveys (distributed in the last weeks of the course), and behavioral data collected in the learning environment. 
% Only those learners who completed the feedback survey were included in the analyses (some did not complete the initial course survey). 
On the initial course surveys, learners reported their gender, age, prior experience, and enrollment intentions (on the OLEI scale \cite{kizilcec2015motivation}). On the feedback survey, learners reported their satisfaction with the course (N-item index, $\alpha=.XX, M=XX, SD=XX$), and whether they `stopped participating in the course before it ended' or `remained active until the end'. If they indicated that they stopped participating, learners reported `what influenced their decision to stop participating' using a grid of 14 binary-choice items (see Figure \ref{fig:s1reason}).\footnote{Instead of a select-all-that-apply question, we opted for binary-choice items to encourage learners to consider each reason in turn. The response options were ``Did influence my decision to stop" and ``Did NOT influence my decision to stop," which should reduce acquiescence bias relative to yes/no or true/false scales.}

@@

[.] What proportion of learners who dropped out were satisfied with their achievements in the course? (RQ2)

\begin{table*}
\label{tab:s1reascor}
\caption{Correlations among reasons for disengagement based on a linear combination of the courses' correlation matrices}
\small
\center
\begin{tabular}{lccccccccccccc}
\toprule
 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 \\
\midrule
1 Assn or exam was too difficult    & 0.17 & 0.28 & 0.30 & 0.16 & 0.14 & 0.37 & 0.24 & 0.26 & 0.20 & 0.09 & 0.33 & 0.40 & 0.28 \\
2 Couldn't keep up with deadlines   & 1 & 0.06 & -0.03 & 0.01 & -0.05 & 0.05 & -0.07 & -0.06 & 0.32 & 0.17 & 0.04 & 0.08 & 0.04 \\
3 Credentials not valuable to me    &   & 1 & 0.34 & 0.34 & 0.25 & 0.34 & 0.37 & 0.39 & 0.17 & 0.17 & 0.27 & 0.28 & 0.35 \\
4 Didn't enjoy online course format &   &  & 1 & 0.24 & 0.15 & 0.53 & 0.36 & 0.44 & 0.02 & 0.18 & 0.34 & 0.33 & 0.54 \\
5 I was only exploring              &   &  &  & 1 & 0.36 & 0.20 & 0.26 & 0.31 & 0.11 & 0.18 & 0.21 & 0.16 & 0.18 \\
6 I learned all I wanted            &   &  &  &  & 1 & 0.12 & 0.31 & 0.28 & 0.0? & 0.12 & 0.16 & 0.16 & 0.09 \\
7 Material not explained properly   &   &  &  &  &  & 1 & 0.36 & 0.45 & 0.15 & 0.13 & 0.31 & 0.35 & 0.66 \\
8 Not challenging enough            &   &  &  &  &  &  & 1 & 0.45 & 0.0 & 0.12 & 0.28 & 0.21 & 0.32 \\
9 Not what I was looking for        &   &  &  &  &  &  &  & 1 & 0.07 & 0.12 & 0.3 & 0.33 & 0.44 \\
10 Course requires too much time    &   &  &  &  &  &  &  &  & 1 & 0.06 & 0.08 & 0.19 & 0.12 \\
11 I joined too late                &   &  &  &  &  &  &  &  &  & 1 & 0.22 & 0.11 & 0.14 \\
12 Technical difficulties           &   &  &  &  &  &  &  &  &  &  & 1 & 0.27 & 0.30 \\
13 Too advanced for my background   &   &  &  &  &  &  &  &  &  &  &  & 1 & 0.33 \\
14 Videos were confusing            &   &  &  &  &  &  &  &  &  &  &  &  & 1 \\
\bottomrule
\end{tabular}
\end{table*}

[.] How do prior learner characteristics (demographics, geographic location, intentions and motivations, and previous online courses) influence learners' likelihood of dropping out of a MOOC? (RQ1)

@@

@@

[.] How do reasons for dropping out vary across prior learner characteristics (cf. RQ1)? (RQ3b) <- we may not be able to answer this

[.] How does behavior in MOOCs vary by reasons for dropping out? (RQ4) <- we need to see if we have enough data to answer this/if this is interesting

>>>>

\begin{abstract}
Research on attrition has a long tradition in the field of education. Large-scale open learning environments prompted a shift away from a monolithic notion of success toward a more nuanced treatment that accounts for diverse motivations. A systematic analysis of predictors of and reported reasons for attrition in 17 massive open online courses suggested time to be a critical obstacle and [other important take-aways from Study 1] (Study 1; $N = XX,XXX$). A follow-up study with online learners who disengaged from the course (Study 2; $ N = 756$) found self-ascribed successful learners to report higher levels of goal striving, feelings of social belonging, and growth mindset. Iterative coding of open responses about challenges learners faced while taking the course reiterated the prevalence of having not enough time (reported by 84\%). A low capacity for volitional control was identified as a large contributor to this issue, besides demands from higher priority obligations. A better understanding of why learners disengage from MOOCs allows for earlier recognition of impending attrition, which enables targeted interventions to help learners achieve their learning goals.
\end{abstract}

@@

Educational environments have become increasingly diverse. Traditional schools and universities have a characteristically rigid structure, including instructor-defined---even nationally agreed---syllabi, fixed time schedules, entry requirements, and material costs to enter and exit. Novel institutional structures have been developed to overcome particular constraints. Community colleges, for instance, were created in an attempt to democratize education by offering instruction at a lower cost, with lower admission criteria, and with more flexible schedules to accommodate those who cannot afford to be full-time students \cite{goldrick2010challenges}. Distance learning programs intended to deliver education in remote parts of the world and for people who simply could not attend in-person classes. Course materials, including assessments, were delivered through mail (correspondence education), radio, television, and eventually the Internet, thereby addressing geographical and time related constraints of traditional instruction \cite{moore1996distance}.

@@

\begin{description}
  \item[RQ1] How do prior learner characteristics (demographics, geographic location, enrollment intentions, and previous online courses) influence learners' likelihood of disengaging from a MOOC?
  \item[H1] Successful MOOC learners exhibit higher levels of goal striving than unsuccessful ones.
\end{description}

@@

\begin{description}
  \item[RQ2a] What proportion of learners who disengaged were satisfied with their achievements in the course?
  \item[RQ2b] How does learners' behavior in the MOOC vary by their level of satisfaction?
  \item[H3] Successful MOOC learners have a more fluid conception of intelligence (i.e., more of a growth than fixed mindset) than unsuccessful ones.
\end{description}

@@

\begin{description}
\item[RQ3a] What reasons do learners report for disengaging from a MOOC?
\item[RQ3b] How do reasons for disengaging vary across prior learner characteristics (cf. RQ1)?
\end{description}

@@

Seventeen MOOCs on topics in a variety of disciplines were selected for investigation (see Table \ref{tab:s1sum}). For each course, we merged data from the initial course survey (distributed in the first weeks of the course), the feedback surveys (distributed in the last weeks of the course), and behavioral data collected in the learning environment. 
% Only those learners who completed the feedback survey were included in the analyses (some did not complete the initial course survey). 
On the initial course surveys, learners reported their gender, age, prior experience, and enrollment intentions (on the OLEI scale \cite{kizilcec2015motivation}). On the feedback survey, learners reported their satisfaction with the course (N-item index, M=XX, SD=XX, $\alpha=.XX$), and whether they `stopped participating in the course before it ended' or `remained active until the end'. If they indicated that they stopped participating, learners reported `what influenced their decision to stop participating' using a grid of 14 binary-choice items (see Figure \ref{fig:s1reason}).\footnote{Instead of a select-all-that-apply question, we opted for binary-choice items to encourage learners to consider each reason in turn. The response options were ``Did influence my decision to stop" and ``Did NOT influence my decision to stop," which should reduce acquiescence bias relative to yes/no or true/false scales.}

@@

Report results of factor analysis here.

@@

Attrition can be a sign of success if a learner has achieved their personal goal. We therefore investigated what proportion of learners who dropped out were satisfied with their achievements in the course (RQ2). TODO add satisfaction analysis here

% \begin{table*}
% \label{tab:s1reascor}
% \caption{Correlations among reasons for disengagement based on a linear combination of the courses' correlation matrices}
% \small
% \center
% \begin{tabular}{lccccccccccccc}
% \toprule
%  & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 \\
% \midrule
% 1 Assn or exam was too difficult    & 0.17 & 0.28 & 0.30 & 0.16 & 0.14 & 0.37 & 0.24 & 0.26 & 0.20 & 0.09 & 0.33 & 0.40 & 0.28 \\
% 2 Couldn't keep up with deadlines   & 1 & 0.06 & -0.03 & 0.01 & -0.05 & 0.05 & -0.07 & -0.06 & 0.32 & 0.17 & 0.04 & 0.08 & 0.04 \\
% 3 Credentials not valuable to me    &   & 1 & 0.34 & 0.34 & 0.25 & 0.34 & 0.37 & 0.39 & 0.17 & 0.17 & 0.27 & 0.28 & 0.35 \\
% 4 Didn't enjoy online course format &   &  & 1 & 0.24 & 0.15 & 0.53 & 0.36 & 0.44 & 0.02 & 0.18 & 0.34 & 0.33 & 0.54 \\
% 5 I was only exploring              &   &  &  & 1 & 0.36 & 0.20 & 0.26 & 0.31 & 0.11 & 0.18 & 0.21 & 0.16 & 0.18 \\
% 6 I learned all I wanted            &   &  &  &  & 1 & 0.12 & 0.31 & 0.28 & 0.0? & 0.12 & 0.16 & 0.16 & 0.09 \\
% 7 Material not explained properly   &   &  &  &  &  & 1 & 0.36 & 0.45 & 0.15 & 0.13 & 0.31 & 0.35 & 0.66 \\
% 8 Not challenging enough            &   &  &  &  &  &  & 1 & 0.45 & 0.0 & 0.12 & 0.28 & 0.21 & 0.32 \\
% 9 Not what I was looking for        &   &  &  &  &  &  &  & 1 & 0.07 & 0.12 & 0.3 & 0.33 & 0.44 \\
% 10 Course requires too much time    &   &  &  &  &  &  &  &  & 1 & 0.06 & 0.08 & 0.19 & 0.12 \\
% 11 I joined too late                &   &  &  &  &  &  &  &  &  & 1 & 0.22 & 0.11 & 0.14 \\
% 12 Technical difficulties           &   &  &  &  &  &  &  &  &  &  & 1 & 0.27 & 0.30 \\
% 13 Too advanced for my background   &   &  &  &  &  &  &  &  &  &  &  & 1 & 0.33 \\
% 14 Videos were confusing            &   &  &  &  &  &  &  &  &  &  &  &  & 1 \\
% \bottomrule
% \end{tabular}
% \end{table*}

% [.] How do prior learner characteristics (demographics, geographic location, intentions and motivations, and previous online courses) influence learners' likelihood of dropping out of a MOOC? (RQ1)

@@
